### 2017-09-12 redaction/online_appendix_dhp2017.tex
Chnagement annexe technique

\documentclass[11pt,twoside, authoryear]{elsarticle}
\usepackage[frenchb, english]{babel}
\usepackage[utf8]{inputenc}
%\usepackage[utf8]{fontenc}
%\usepackage[ansinew]{inputenc}
\usepackage{setspace}
\usepackage{babel,varioref}
\usepackage{supertabular}
\usepackage{graphicx}
%\usepackage[pdftex]{graphicx}
%\DeclareGraphicsExtensions{.pdf,.mps,.png,.jpg,.eps}
\usepackage{lscape}
\usepackage[authoryear]{natbib}
%\usepackage[frenchb,english]{babel}
\usepackage{multirow}
\usepackage{ifthen}
%\usepackage{harvard}
\usepackage{vmargin}
\usepackage{verbatim}
\usepackage{array}
%\usepackage[authoryear]{natbib}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=Black, citecolor=blue, urlcolor=blue}
%\setmarginsrb{2cm}{2cm}{2cm}{2cm}{0cm}{0cm}{0cm}{0cm}
\usepackage{booktabs,caption,fixltx2e}

\usepackage[flushleft]{threeparttable}

%\usepackage{titling}

%\setlength{\droptitle}{-10em}   % This is your set screw

\makeatletter
\def\ps@pprintTitle{%
  \let\@oddhead\@empty
  \let\@evenhead\@empty
  \let\@oddfoot\@empty
  \let\@evenfoot\@oddfoot
}
\makeatother
\usepackage{adjustbox}
\usepackage{chngcntr}
\counterwithin{table}{section}
%\usepackage[round]{natbib}

\renewcommand{\thesection}{\Alph{section}}

% \newcommand\section[1]{%
%   \refstepcounter{section}%
%   \addcontentsline{toc}{section}{\protect\numberline{\thesection}#1}%
%   \sectionmark{#1}}
%   }

% % \refstepcounter{section}%
% %   \addcontentsline{toc}{section}{\protect\numberline{\thesection}#1}%
% %   \sectionmark{#1}}

%   \newcommand\subsection[1]{%
%   \refstepcounter{subsection}%
%   \addcontentsline{toc}{subsection}{\protect\numberline{\thesubsection}#1}%
%   %\subsectionmark{#1}
%   }

%\EnableSectionsInLOFT

\begin{document}
\title{\textbf{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}\\Online Appendix}

\urlstyle{rm}
\author{Guillaume Daudin\footnote{\noindent Corresponding author. Universit\'{e} Paris-Dauphine, PSL Research University, LEDa-DIAL, UMR 225, France \&
SciencesPo, Observatoire Français des Conjonctures \'{E}conomiques (OFCE), France; email: \url{guillaume.daudin@dauphine.fr}} \qquad \quad Jérôme Héricourt\footnote{\noindent Corresponding author. Université de Lille - LEM-CNRS (UMR 9221) and CEPII; email: \url{jerome.hericourt@univ-lille1.fr}} \qquad \quad Lise Patureau\footnote{\noindent Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, France ; email: \url{lise.patureau@dauphine.fr}}}
%\date{\vspace{-5ex}}

%Modèle additif seul DONE
%+Tableaux avec toutes les années en 3 digit DONE
% Tableaux avec toutes les années en 4 digit ???
%Harmoniser la présentation des résultats (cf table 1 du papier) DONE
%Les résultats sur la version Hummels mais avec sa pondération: en annexe du papier? En online appendix? DONE, mais à compléter

\maketitle

%\newpage

\tableofcontents
\vspace{1cm}
\newpage
\listoftables

\newpage

%
%
%
%
%			SECTION A / ADDITIVE ONLY
%			
%			
%				
%
%

	\renewcommand\thesubsubsection{\Alph{subsection}.\arabic{subsubsection}}

	\renewcommand\thesubsection{\Alph{subsection}}

\subsection{Three models: Comparison \label{secoa:additive_only}}

%%%% RAW TABLE TO BE FOUND AT: "...\Dropbox\Papier_Lise_Guillaume\trade_cost\results\3_models"
% modele iceberg tout seul : terme nlI
% colonnes L à U: modèle avec les deux
% termes nla, de V à Z, modèles avec que de l'additif
%nlI quer iceberg
%nla que additif
%nl : les deux
\setcounter{table}{0}
\renewcommand{\thetable}{A.\arabic{table}}
%\subsubsection{Nominal Exchange Rate Volatility}

\begin{table}[htbp]
  \centering
    \caption{Estimation results of the three models (Air, 3-digit level)}
  \label{tab:3models_estimation_results_air}%
    \begin{tabular}{l|c|c|c|c|c|c|c|c|c}
    \hline \hline
    \textbf{Year} & \multicolumn{1}{c}{\textbf{1974}} & \multicolumn{1}{c}{\textbf{1980}} & \multicolumn{1}{c}{\textbf{1985}} & \multicolumn{1}{c}{\textbf{1990}} & \multicolumn{1}{c}{\textbf{1995}} & \multicolumn{1}{c}{\textbf{2000}} & \multicolumn{1}{c}{\textbf{2005}} & \multicolumn{1}{c}{\textbf{2010}} & \multicolumn{1}{c}{\textbf{2013}} \\ \hline
   $\#$ obs. & \multicolumn{1}{c}{14955} & \multicolumn{1}{c}{16118} & \multicolumn{1}{c}{19908} & \multicolumn{1}{c}{24958} & \multicolumn{1}{c}{31037} & \multicolumn{1}{c}{35027} & \multicolumn{1}{c}{41806} & \multicolumn{1}{c}{40279} & \multicolumn{1}{c}{39351} \\
   $\#$  origin countries & \multicolumn{1}{c}{152} & \multicolumn{1}{c}{165} & \multicolumn{1}{c}{169} & \multicolumn{1}{c}{181} & \multicolumn{1}{c}{207} & \multicolumn{1}{c}{208} & \multicolumn{1}{c}{211} & \multicolumn{1}{c}{210} & \multicolumn{1}{c}{210} \\
   $\#$  sectors & \multicolumn{1}{c}{203} & \multicolumn{1}{c}{204} & \multicolumn{1}{c}{207} & \multicolumn{1}{c}{212} & \multicolumn{1}{c}{217} & \multicolumn{1}{c}{218} & \multicolumn{1}{c}{217} & \multicolumn{1}{c}{216} & \multicolumn{1}{c}{212} \\ \hline
    \multicolumn{10}{l}{\textbf{Model (A) - Iceberg transport costs only} ($\widehat{\tau}^{ice}$)} \\ \hline
    Mean (in \%) & \multicolumn{1}{c}{6.93} & \multicolumn{1}{c}{5.41} & \multicolumn{1}{c}{6.08} & \multicolumn{1}{c}{5.03} & \multicolumn{1}{c}{4.61} & \multicolumn{1}{c}{3.60} & \multicolumn{1}{c}{4.10} & \multicolumn{1}{c}{4.19} & \multicolumn{1}{c}{3.36} \\
    Median (in \%)& \multicolumn{1}{c}{5.43} & \multicolumn{1}{c}{3.79} & \multicolumn{1}{c}{5.47} & \multicolumn{1}{c}{4.37} & \multicolumn{1}{c}{3.80} & \multicolumn{1}{c}{2.47} & \multicolumn{1}{c}{3.12} & \multicolumn{1}{c}{3.41} & \multicolumn{1}{c}{2.92} \\
    Std & \multicolumn{1}{c}{0.04} & \multicolumn{1}{c}{0.04} & \multicolumn{1}{c}{0.04} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} \\ \hline
    \multicolumn{10}{l}{\textbf{Model (B) - With additive and ad-valorem transport costs}} \\ \hline
     \multicolumn{10}{l}{ \textit{Multiplicative term} ($\widehat{\tau}^{adv}$)} \\ \hline
    Mean (in \%) & \multicolumn{1}{c}{3.64} & \multicolumn{1}{c}{2.32} & \multicolumn{1}{c}{2.46} & \multicolumn{1}{c}{2.38} & \multicolumn{1}{c}{2.05} & \multicolumn{1}{c}{1.66} & \multicolumn{1}{c}{2.00} & \multicolumn{1}{c}{2.57} & \multicolumn{1}{c}{1.70} \\
    Median (in \%)& \multicolumn{1}{c}{2.71} & \multicolumn{1}{c}{1.57} & \multicolumn{1}{c}{1.79} & \multicolumn{1}{c}{1.60} & \multicolumn{1}{c}{1.39} & \multicolumn{1}{c}{1.20} & \multicolumn{1}{c}{1.57} & \multicolumn{1}{c}{2.24} & \multicolumn{1}{c}{1.72} \\
    Std & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.01} \\ \hline
      \multicolumn{10}{l}{\textit{Additive term} ($\widehat{t}/\widetilde{p}$)} \\ \hline
    Mean (in \%) & \multicolumn{1}{c}{2.56} & \multicolumn{1}{c}{2.04} & \multicolumn{1}{c}{2.83} & \multicolumn{1}{c}{1.83} & \multicolumn{1}{c}{1.64} & \multicolumn{1}{c}{1.30} & \multicolumn{1}{c}{1.43} & \multicolumn{1}{c}{1.13} & \multicolumn{1}{c}{1.01} \\
    Median (in \%)& \multicolumn{1}{c}{1.13} & \multicolumn{1}{c}{0.54} & \multicolumn{1}{c}{1.30} & \multicolumn{1}{c}{0.84} & \multicolumn{1}{c}{0.68} & \multicolumn{1}{c}{0.45} & \multicolumn{1}{c}{0.53} & \multicolumn{1}{c}{0.43} & \multicolumn{1}{c}{0.47} \\
    Std & \multicolumn{1}{c}{0.04} & \multicolumn{1}{c}{0.04} & \multicolumn{1}{c}{0.04} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} \\ \hline
    \multicolumn{10}{l}{\textbf{Model (C) - With additive transport costs only} ($\widehat{t}^{add}/\widetilde{p}$)} \\ \hline
    Mean (in \%) & \multicolumn{1}{c}{6.88} & \multicolumn{1}{c}{4.82} & \multicolumn{1}{c}{5.99} & \multicolumn{1}{c}{4.44} & \multicolumn{1}{c}{3.80} & \multicolumn{1}{c}{3.09} & \multicolumn{1}{c}{4.00} & \multicolumn{1}{c}{4.49} & \multicolumn{1}{c}{3.29} \\
    Median (in \%)& \multicolumn{1}{c}{4.45} & \multicolumn{1}{c}{1.82} & \multicolumn{1}{c}{3.36} & \multicolumn{1}{c}{2.27} & \multicolumn{1}{c}{1.72} & \multicolumn{1}{c}{1.39} & \multicolumn{1}{c}{1.93} & \multicolumn{1}{c}{2.72} & \multicolumn{1}{c}{2.06} \\
    Std & \multicolumn{1}{c}{0.09} & \multicolumn{1}{c}{0.08} & \multicolumn{1}{c}{0.08} & \multicolumn{1}{c}{0.10} & \multicolumn{1}{c}{0.09} & \multicolumn{1}{c}{0.06} & \multicolumn{1}{c}{0.07} & \multicolumn{1}{c}{0.07} & \multicolumn{1}{c}{0.05} \\  \hline \hline
    \end{tabular}
\end{table}%

\begin{landscape}
\begin{table}[htbp]
  \centering
  \footnotesize{
  \caption{Quality-of-fit diagnostic tests of the three models (Air, 3-digit level)}
    \label{tab:3models_diagnosis_air}%
   \begin{tabular}{l|c|c|c|c|c|c|c|c|c}
    \hline \hline
    \textbf{Year} & \multicolumn{1}{c}{\textbf{1974}} & \multicolumn{1}{c}{\textbf{1980}} & \multicolumn{1}{c}{\textbf{1985}} & \multicolumn{1}{c}{\textbf{1990}} & \multicolumn{1}{c}{\textbf{1995}} & \multicolumn{1}{c}{\textbf{2000}} & \multicolumn{1}{c}{\textbf{2005}} & \multicolumn{1}{c}{\textbf{2010}} & \multicolumn{1}{c}{\textbf{2013}} \\ \hline
    \multicolumn{10}{l}{\textbf{$R^2$}} \\ \hline
    Model (A) & \multicolumn{1}{c}{0.297} & \multicolumn{1}{c}{0.267} & \multicolumn{1}{c}{0.302} & \multicolumn{1}{c}{0.251} & \multicolumn{1}{c}{0.142} & \multicolumn{1}{c}{0.318} & \multicolumn{1}{c}{0.460} & \multicolumn{1}{c}{0.421} & \multicolumn{1}{c}{0.313} \\
    Model (B ) & \multicolumn{1}{c}{0.594} & \multicolumn{1}{c}{0.646} & \multicolumn{1}{c}{0.635} & \multicolumn{1}{c}{0.627} & \multicolumn{1}{c}{0.658} & \multicolumn{1}{c}{0.640} & \multicolumn{1}{c}{0.593} & \multicolumn{1}{c}{0.513} & \multicolumn{1}{c}{0.419} \\
    Model (C) & \multicolumn{1}{c}{0.489} & \multicolumn{1}{c}{0.543} & \multicolumn{1}{c}{0.531} & \multicolumn{1}{c}{0.517} & \multicolumn{1}{c}{0.546} & \multicolumn{1}{c}{0.518} & \multicolumn{1}{c}{0.464} & \multicolumn{1}{c}{0.339} & \multicolumn{1}{c}{0.295} \\ \hline
    \multicolumn{10}{l}{\textbf{SER}}\\ \hline
    Model (A) & \multicolumn{1}{c}{0.791} & \multicolumn{1}{c}{0.860} & \multicolumn{1}{c}{0.831} & \multicolumn{1}{c}{0.811} & \multicolumn{1}{c}{0.798} & \multicolumn{1}{c}{0.844} & \multicolumn{1}{c}{0.837} & \multicolumn{1}{c}{0.857} & \multicolumn{1}{c}{0.920} \\
    Model (B) & \multicolumn{1}{c}{0.674} & \multicolumn{1}{c}{0.715} & \multicolumn{1}{c}{0.692} & \multicolumn{1}{c}{0.675} & \multicolumn{1}{c}{0.641} & \multicolumn{1}{c}{0.697} & \multicolumn{1}{c}{0.727} & \multicolumn{1}{c}{0.787} & \multicolumn{1}{c}{0.847} \\
    Model (C) & \multicolumn{1}{c}{1.610} & \multicolumn{1}{c}{1.778} & \multicolumn{1}{c}{1.736} & \multicolumn{1}{c}{1.699} & \multicolumn{1}{c}{1.700} & \multicolumn{1}{c}{1.786} & \multicolumn{1}{c}{1.783} & \multicolumn{1}{c}{1.776} & \multicolumn{1}{c}{1.723} \\ \hline
   \multicolumn{10}{l}{ \textbf{AIC criteria}} \\ \hline
    Model (A) & \multicolumn{1}{c}{35675.0} & \multicolumn{1}{c}{41171.0} & \multicolumn{1}{c}{49315.0} & \multicolumn{1}{c}{60715.6} & \multicolumn{1}{c}{74386.4} & \multicolumn{1}{c}{87492.5} & \multicolumn{1}{c}{103983.0} & \multicolumn{1}{c}{102297.7} & \multicolumn{1}{c}{106130.6} \\
    Model (B ) & \multicolumn{1}{c}{31387.3} & \multicolumn{1}{c}{35738.4} & \multicolumn{1}{c}{42535.8} & \multicolumn{1}{c}{52098.9} & \multicolumn{1}{c}{61343.7} & \multicolumn{1}{c}{74954.9} & \multicolumn{1}{c}{92758.6} & \multicolumn{1}{c}{95887.1} & \multicolumn{1}{c}{100155.4} \\
    Model (C) & \multicolumn{1}{c}{40808.1} & \multicolumn{1}{c}{45138.5} & \multicolumn{1}{c}{55214.8} & \multicolumn{1}{c}{69458.5} & \multicolumn{1}{c}{83958.6} & \multicolumn{1}{c}{100040.8} & \multicolumn{1}{c}{123592.1} & \multicolumn{1}{c}{129359.0} & \multicolumn{1}{c}{127399.2} \\ \hline
    \multicolumn{10}{l}{\textbf{Log-likelihood}}\\ \hline
    Model (A) & \multicolumn{1}{c}{-17530.5} & \multicolumn{1}{c}{-20253.5} & \multicolumn{1}{c}{-24315.5} & \multicolumn{1}{c}{-29977.8} & \multicolumn{1}{c}{-36811.2} & \multicolumn{1}{c}{-43341.3} & \multicolumn{1}{c}{-51648.5} & \multicolumn{1}{c}{-50746.8} & \multicolumn{1}{c}{-52690.3} \\
    Model (B ) & \multicolumn{1}{c}{-15125.6} & \multicolumn{1}{c}{-17263.2} & \multicolumn{1}{c}{-20686.9} & \multicolumn{1}{c}{-25393.5} & \multicolumn{1}{c}{-30036.9} & \multicolumn{1}{c}{-36788.4} & \multicolumn{1}{c}{-45768.3} & \multicolumn{1}{c}{-47277.5} & \multicolumn{1}{c}{-49419.7} \\
    Model (C) & \multicolumn{1}{c}{-20074.1} & \multicolumn{1}{c}{-22217.2} & \multicolumn{1}{c}{-27251.4} & \multicolumn{1}{c}{-34355.3} & \multicolumn{1}{c}{-41634.3} & \multicolumn{1}{c}{-49625.4} & \multicolumn{1}{c}{-61533.0} & \multicolumn{1}{c}{-64339.5} & \multicolumn{1}{c}{-63316.6} \\ \hline
    \multicolumn{10}{l}{\textbf{Test LL} }\\ \hline
    Stat LL ratio (B vs A) & \multicolumn{1}{c}{4809.7} & \multicolumn{1}{c}{5980.6} & \multicolumn{1}{c}{7257.3} & \multicolumn{1}{c}{9168.7} & \multicolumn{1}{c}{13548.7} & \multicolumn{1}{c}{13105.7} & \multicolumn{1}{c}{11760.4} & \multicolumn{1}{c}{6938.6} & \multicolumn{1}{c}{6541.2} \\
    \textbackslash{}\# restrictions & \multicolumn{1}{c}{16929} & \multicolumn{1}{c}{18098} & \multicolumn{1}{c}{21893} & \multicolumn{1}{c}{26948} & \multicolumn{1}{c}{33032} & \multicolumn{1}{c}{37027} & \multicolumn{1}{c}{43811} & \multicolumn{1}{c}{42289} & \multicolumn{1}{c}{41364} \\
    p-value & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} \\ \hline
    Stat LL ratio (B vs C) & \multicolumn{1}{c}{4948.4} & \multicolumn{1}{c}{4954.0} & \multicolumn{1}{c}{6564.5} & \multicolumn{1}{c}{8961.8} & \multicolumn{1}{c}{11597.4} & \multicolumn{1}{c}{12837.0} & \multicolumn{1}{c}{15764.7} & \multicolumn{1}{c}{17062.0} & \multicolumn{1}{c}{13896.9} \\
    \textbackslash{}\# restrictions & \multicolumn{1}{c}{16929} & \multicolumn{1}{c}{18098} & \multicolumn{1}{c}{21893} & \multicolumn{1}{c}{26948} & \multicolumn{1}{c}{33032} & \multicolumn{1}{c}{37027} & \multicolumn{1}{c}{43811} & \multicolumn{1}{c}{42289} & \multicolumn{1}{c}{41364} \\
    p-value & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} \\ \hline \hline
\end{tabular}%
}
\end{table}%
\end{landscape}

\begin{table}[htbp]
  \centering
    \caption{Estimation results of the three models (Vessel, 3-digit level)}
  \label{tab:3models_estimation_results_vessel}%
    \begin{tabular}{l|c|c|c|c|c|c|c|c|c}
    \hline \hline
    \textbf{Year} & 1974  & 1980  & 1985  & 1990  & 1995  & 2000  & 2005  & 2010  & 2013 \\ \hline
    $\# $obs. & 19007 & 17356 & 23348 & 28383 & 32146 & 36090 & 41319 & 37748 & 38473 \\
    $\#$  origin countries & 154   & 163   & 171   & 179   & 201   & 206   & 206   & 198   & 203 \\
    $\#$  sectors & 239   & 232   & 232   & 232   & 228   & 230   & 231   & 226   & 224 \\ \hline
    \multicolumn{10}{l}{\textbf{Model (A) - Iceberg transport costs only} ($\widehat{\tau}^{ice}$)} \\ \hline
    Mean (in \%) & 9.79  & 6.53  & 6.88  & 5.67  & 5.14  & 5.10  & 5.47  & 3.99  & 3.60 \\
    Median (in \%)& 9.58  & 5.50  & 6.33  & 4.63  & 4.29  & 4.85  & 4.90  & 3.56  & 3.28 \\
    Std & 0.05  & 0.04  & 0.04  & 0.03  & 0.03  & 0.03  & 0.03  & 0.02  & 0.02 \\ \hline
    \multicolumn{10}{l}{\textbf{Model (B ) - With additive and ad-valorem transport costs} } \\ \hline
     \multicolumn{10}{l}{\textit{Multiplicative term} ($\widehat{\tau}^{adv}$)} \\ \hline
    Mean (in \%) & 5.42  & 3.08  & 4.02  & 3.31  & 2.79  & 2.49  & 2.68  & 1.95  & 2.22 \\
    Median (in \%) & 4.93  & 2.42  & 3.60  & 2.81  & 2.53  & 2.07  & 2.08  & 1.76  & 1.82 \\
    Std & 0.04  & 0.02  & 0.03  & 0.02  & 0.02  & 0.02  & 0.02  & 0.02  & 0.01 \\ \hline
       \multicolumn{10}{l}{\textit{Additive term} ($\widehat{t}/\widetilde{p}$)}  \\ \hline
    Mean (in \%)  & 5.08  & 3.38  & 3.19  & 2.73  & 2.73  & 2.80  & 3.02  & 2.47  & 1.46 \\
    Median (in \%) & 2.94  & 2.27  & 2.06  & 1.70  & 1.82  & 2.19  & 2.16  & 1.89  & 0.76 \\
    Std & 0.09  & 0.05  & 0.04  & 0.04  & 0.04  & 0.04  & 0.04  & 0.03  & 0.02 \\ \hline
    \multicolumn{10}{l}{\textbf{Model (C) - With additive transport costs only} ($\widehat{t}^{add}/\widetilde{p}$)} \\ \hline
    Mean (in \%) & 14.51 & 10.05 & 10.47 & 14.62 & 8.37  & 8.02  & 8.41  & 6.40  & 5.23 \\
    Median (in \%)& 9.53  & 6.69  & 7.16  & 6.22  & 4.68  & 4.93  & 5.74  & 3.87  & 3.57 \\
    Std & 0.24  & 0.17  & 0.18  & 0.30  & 0.15  & 0.16  & 0.15  & 0.15  & 0.10 \\ \hline \hline
   \end{tabular}%
\end{table}%

\begin{landscape}
\begin{table}[htbp]
  \centering
  \footnotesize{
  \caption{Quality-of-fit diagnostic tests of the three models (Vessel, 3-digit level)}
    \label{tab:3models_diagnosis_air}%
   \begin{tabular}{l|c|c|c|c|c|c|c|c|c}
    \hline \hline
    \textbf{Year} & \multicolumn{1}{c}{\textbf{1974}} & \multicolumn{1}{c}{\textbf{1980}} & \multicolumn{1}{c}{\textbf{1985}} & \multicolumn{1}{c}{\textbf{1990}} & \multicolumn{1}{c}{\textbf{1995}} & \multicolumn{1}{c}{\textbf{2000}} & \multicolumn{1}{c}{\textbf{2005}} & \multicolumn{1}{c}{\textbf{2010}} & \multicolumn{1}{c}{\textbf{2013}} \\ \hline
    \multicolumn{10}{l}{\textbf{$R^2$}} \\ \hline
    Model (A) & 0,450 & 0,415 & 0,427 & 0,456 & 0,438 & 0,401 & 0,378 & 0,350 & 0,339 \\
    Model (B) & 0,612 & 0,575 & 0,571 & 0,590 & 0,611 & 0,571 & 0,541 & 0,491 & 0,462 \\
    Model (C ) & 0,424 & 0,401 & 0,374 & 0,429 & 0,456 & 0,431 & 0,417 & 0,358 & 0,349 \\ \hline
    \multicolumn{10}{l}{\textbf{SER} } \\ \hline
    Model (A) & 0,576 & 0,620 & 0,569 & 0,592 & 0,615 & 0,652 & 0,673 & 0,740 & 0,758 \\
    Model (B) & 0,484 & 0,528 & 0,493 & 0,514 & 0,512 & 0,551 & 0,578 & 0,656 & 0,684 \\
    Model (C ) & 1,271 & 1,339 & 1,283 & 1,326 & 1,302 & 1,319 & 1,336 & 1,392 & 1,410 \\ \hline
    \multicolumn{10}{l}{\textbf{AIC criteria} } \\ \hline
    Model (A) & 33328,81 & 33010,27 & 40275,70 & 51142,62 & 60414,92 & 71365,89 & 85051,02 & 84789,89 & 88191,87 \\
    Model (B) & 27331,52 & 28067,31 & 34170,52 & 43664,74 & 49275,33 & 60475,91 & 73020,09 & 76161,33 & 80873,72 \\
    Model (C ) & 46082,40 & 44370,26 & 58829,71 & 71461,52 & 77052,41 & 88746,51 & 103310,93 & 101166,91 & 104290,27 \\ \hline
    \multicolumn{10}{l}{\textbf{Log-likelihood}}\\ \hline
    Model (A) & -16287,40 & -16129,13 & -19767,85 & -25169,31 & -29790,46 & -35263,95 & -42122,51 & -41998,95 & -43692,93 \\
    Model (B) & -12985,76 & -13353,65 & -16398,26 & -21171,37 & -23905,66 & -29490,96 & -35844,04 & -37418,66 & -39751,86 \\
    Model (C ) & -22674,20 & -21814,13 & -29045,86 & -35403,76 & -38125,20 & -43963,25 & -51245,46 & -50348,45 & -51783,14 \\ \hline
   \multicolumn{10}{l}{ \textbf{Test LL} } \\ \hline
    Stat LL ratio (B vs A) & 6603,28 & 5550,96 & 6739,18 & 7995,88 & 11769,59 & 11545,98 & 12556,94 & 9160,56 & 7882,15 \\
    $\#$ restrictions & 393   & 395   & 403   & 411   & 429   & 436   & 437   & 424   & 427 \\
    p-value & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00 \\ \hline
    Stat LL ratio (B vs C) & 19376,88 & 16920,95 & 25295,20 & 28464,79 & 28439,08 & 28944,59 & 30802,84 & 25859,58 & 24062,55 \\
    $\#$ restrictions & 393   & 395   & 403   & 411   & 429   & 436   & 437   & 424   & 427 \\
    p-value & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00 \\
\hline \hline
    \end{tabular}%
}
\end{table}%
\end{landscape}

%\subsection{Detailed Results \label{secoa:detailed}}
\setcounter{table}{0}
\renewcommand{\thetable}{B.\arabic{table}}

\subsection{Transport Cost Estimates: Yearly Detailed Results}

\subsubsection{3-Digits Level Product Classification}
\begin{landscape}
\begin{table}[htbp]
\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
  %\centering
\caption{Air: Transport costs estimates, all years, 3-digit}
\begin{center}
\scalebox{0.85}{\begin{tabular}{lcccccccccccccccccccc}
\hline
\hline
Year  & 1974  & 1975  & 1976  & 1977  & 1978  & 1979  & 1980  & 1981  & 1982  & 1983  & 1984  & 1985  & 1986  & 1987  & 1988  & 1989  & 1990  & 1991  & 1992  & 1993 \\
\hline
\multicolumn{20}{l}{\textbf{Model (A) - With only Ad-Valorem Trade Costs} ($\widehat{\tau}^{ice}$, in \%)} \\
\hline
Mean  & 6.9   & 7.5   & 7.2   & 7.7   & 6.9   & 6.1   & 5.4   & 6.0   & 6.4   & 6.9   & 7.2   & 6.1   & 6.4   & 6.6   & 5.7   & 5.3   & 5.0   & 5.1   & 4.9   & 5.1 \\
Median & 5.4   & 6.4   & 6.9   & 7.1   & 6.3   & 5.3   & 3.8   & 4.9   & 5.3   & 6.1   & 6.7   & 5.5   & 5.9   & 6.3   & 5.3   & 4.6   & 4.4   & 4.5   & 4.5   & 4.4 \\
\hline
\multicolumn{20}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem Trade Costs} }\\ \hline
\multicolumn{20}{l}{\textit{Ad-valorem term ($\widehat{\tau}^{adv}$, in \%)} }   \\
\hline
Mean  & 3.6   & 3.7   & 3.9   & 3.8   & 3.2   & 3.0   & 2.3   & 2.8   & 2.8   & 2.6   & 3.3   & 2.5   & 3.2   & 2.6   & 3.1   & 3.1   & 2.4   & 2.7   & 2.2   & 2.4 \\
Median & 2.7   & 2.7   & 2.9   & 2.7   & 2.1   & 2.4   & 1.6   & 1.8   & 1.9   & 1.9   & 2.7   & 1.8   & 2.1   & 2.0   & 2.0   & 1.9   & 1.6   & 1.5   & 1.5   & 1.6 \\

\hline
\multicolumn{20}{l}{\textit{Additive term ($\widehat{t}^{add}/\widetilde{p}$, in \%)} }   \\
\hline
Mean  & 2.6   & 3.0   & 2.3   & 3.1   & 2.6   & 2.1   & 2.0   & 2.0   & 2.3   & 2.8   & 2.5   & 2.8   & 2.6   & 2.9   & 1.7   & 4.6   & 1.8   & 1.8   & 1.9   & 1.9 \\
Median & 1.1   & 1.2   & 0.9   & 1.3   & 1.1   & 0.7   & 0.5   & 0.6   & 0.8   & 1.0   & 1.0   & 1.3   & 1.3   & 1.5   & 1.0   & 0.7   & 0.8   & 0.6   & 0.9   & 0.8 \\
\hline
\# observations & 14955 & 15299 & 11397 & 10707 & 15222 & 15684 & 16118 & 16864 & 17322 & 18180 & 20644 & 19908 & 20695 & 20793 & 24663 & 25197 & 24958 & 25156 & 26191 & 28296 \\

%%%%%%%%%%%%%%%%2nd part of the table

\hline\hline
\multicolumn{20}{c}{ }  \\
\multicolumn{20}{c}{Continued}  \\
\hline\hline

Year  & 1994  & 1995  & 1996  & 1997  & 1998  & 1999  & 2000  & 2001  & 2002  & 2003  & 2004  & 2005  & 2006  & 2007  & 2008  & 2009  & 2010  & 2011  & 2012  & 2013 \\
\hline
\multicolumn{20}{l}{\textbf{Model (A) - With only Ad-Valorem Trade Costs} ($\widehat{\tau}^{ice}$, in \%)} \\
\hline
Mean  & 4.6   & 4.6   & 4.2   & 4.1   & 3.8   & 3.8   & 3.6   & 3.5   & 3.8   & 3.9   & 4.0   & 4.1   & 3.9   & 4.1   & 4.1   & 4.0   & 4.2   & 3.9   & 3.7   & 3.4 \\
Median & 3.7   & 3.8   & 3.1   & 3.0   & 2.7   & 2.8   & 2.5   & 2.4   & 2.7   & 2.6   & 2.9   & 3.1   & 2.7   & 3.0   & 3.2   & 3.0   & 3.4   & 3.1   & 3.0   & 2.9 \\
\hline
\multicolumn{20}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem Trade Costs} }\\ \hline
\multicolumn{20}{l}{\textit{Ad-valorem term ($\widehat{\tau}^{adv}$, in \%)} }   \\
\hline
Mean  & 2.3   & 2.1   & 1.9   & 1.8   & 1.8   & 1.8   & 1.7   & 1.6   & 1.6   & 1.9   & 1.9   & 2.0   & 1.8   & 2.3   & 2.3   & 2.3   & 2.6   & 2.2   & 2.2   & 1.7 \\
Median & 1.3   & 1.4   & 1.4   & 1.3   & 1.3   & 1.5   & 1.2   & 1.1   & 1.2   & 1.4   & 1.4   & 1.6   & 1.4   & 1.9   & 1.9   & 1.8   & 2.2   & 1.7   & 1.9   & 1.7 \\
\hline
\multicolumn{20}{l}{\textit{Additive term ($\widehat{t}^{add}/\widetilde{p}$, in \%)} }   \\
\hline
Mean  & 1.7   & 1.6   & 1.5   & 1.5   & 1.4   & 1.4   & 1.3   & 1.3   & 1.6   & 1.4   & 1.5   & 1.4   & 1.3   & 1.2   & 1.2   & 1.2   & 1.1   & 1.1   & 0.9   & 1.0 \\
Median & 0.8   & 0.7   & 0.6   & 0.6   & 0.5   & 0.5   & 0.5   & 0.5   & 0.5   & 0.5   & 0.6   & 0.5   & 0.5   & 0.5   & 0.5   & 0.5   & 0.4   & 0.4   & 0.4   & 0.5 \\
\hline
\# observations & 29948 & 31037 & 32187 & 33502 & 33492 & 33523 & 35027 & 34885 & 35159 & 35891 & 36990 & 41806 & 42554 & 40858 & 40159 & 38275 & 40279 & 41190 & 40909 & 39351 \\

\hline
\hline
\end{tabular}}%

\end{center}
\label{tab_oa:result_air_ally3}%
\end{table}%
\end{landscape}

\begin{landscape}
\begin{table}[htbp]
\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
\caption{Vessel: Transport costs estimates, all years, 3-digit}
\begin{center}
\scalebox{0.85}{\begin{tabular}{lcccccccccccccccccccc}
\hline
\hline
Year  & 1974  & 1975  & 1976  & 1977  & 1978  & 1979  & 1980  & 1981  & 1982  & 1983  & 1984  & 1985  & 1986  & 1987  & 1988  & 1989  & 1990  & 1991  & 1992  & 1993 \\
\hline
\multicolumn{20}{l}{\textbf{Model (A) - With only Ad-Valorem Trade Costs} ($\widehat{\tau}^{ice}$, in \%)} \\
\hline
Mean  & 9.8   & 9.9   & 8.9   & 8.3   & 8.1   & 7.5   & 6.5   & 6.0   & 6.3   & 7.0   & 7.0   & 6.9   & 6.7   & 6.2   & 6.1   & 5.7   & 5.7   & 5.5   & 5.0   & 5.2 \\
Median & 9.6   & 8.5   & 8.0   & 7.3   & 7.1   & 6.5   & 5.5   & 5.0   & 5.9   & 5.7   & 6.1   & 6.3   & 7.0   & 6.3   & 5.7   & 4.8   & 4.6   & 4.4   & 4.2   & 4.7 \\
\hline
\multicolumn{20}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem Trade Costs} }\\ \hline
\multicolumn{20}{l}{\textit{Ad-valorem term ($\widehat{\tau}^{adv}$, in \%)} }   \\
\hline
Mean  & 5.4   & 4.8   & 5.4   & 5.2   & 5.9   & 4.6   & 3.1   & 3.3   & 3.4   & 4.2   & 4.1   & 4.0   & 3.9   & 3.6   & 4.0   & 3.0   & 3.3   & 3.0   & 2.6   & 2.9 \\
Median & 4.9   & 4.1   & 4.8   & 4.4   & 5.4   & 4.0   & 2.4   & 2.9   & 2.9   & 3.9   & 3.5   & 3.6   & 3.6   & 3.0   & 3.5   & 2.6   & 2.8   & 2.7   & 2.3   & 2.6 \\
\multicolumn{20}{l}{\textit{Additive term ($\widehat{t}^{add}/\widetilde{p}$, in \%)} }   \\
\hline
Mean  & 5.1   & 5.5   & 3.5   & 3.5   & 2.5   & 3.1   & 3.4   & 2.9   & 3.5   & 2.9   & 3.2   & 3.2   & 2.9   & 2.8   & 2.4   & 2.9   & 2.7   & 2.8   & 2.7   & 2.7 \\
Median & 2.9   & 3.6   & 1.9   & 1.7   & 1.2   & 1.7   & 2.3   & 1.5   & 2.3   & 2.0   & 2.3   & 2.1   & 1.8   & 1.8   & 1.3   & 2.0   & 1.7   & 1.7   & 1.8   & 1.6 \\
\hline
\# observations & 19007 & 18710 & 13615 & 12826 & 16601 & 17274 & 17356 & 17788 & 18075 & 18883 & 21650 & 23348 & 23729 & 23626 & 27661 & 29106 & 28383 & 28095 & 29050 & 30839 \\ %%%%%%%%%%%%%%%%2nd part of the table

\hline\hline
\multicolumn{20}{c}{ }  \\
\multicolumn{20}{c}{Continued}  \\
\hline\hline

Year  & 1994  & 1995  & 1996  & 1997  & 1998  & 1999  & 2000  & 2001  & 2002  & 2003  & 2004  & 2005  & 2006  & 2007  & 2008  & 2009  & 2010  & 2011  & 2012  & 2013 \\
\hline
\multicolumn{20}{l}{\textbf{Model (A) - With only Ad-Valorem Trade Costs} ($\widehat{\tau}^{ice}$, in \%)} \\
\hline
Mean  & 5.2   & 5.1   & 4.8   & 4.7   & 4.8   & 5.0   & 5.1   & 5.0   & 4.8   & 5.2   & 5.4   & 5.5   & 4.8   & 4.7   & 4.4   & 4.3   & 4.0   & 3.5   & 3.6   & 3.6 \\
Median & 4.1   & 4.3   & 3.9   & 3.9   & 3.9   & 4.5   & 4.9   & 4.6   & 4.1   & 4.8   & 5.1   & 4.9   & 4.2   & 4.2   & 3.8   & 4.1   & 3.6   & 3.0   & 3.1   & 3.3 \\
\hline
\multicolumn{20}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem Trade Costs} }\\ \hline
\multicolumn{20}{l}{\textit{Ad-valorem term ($\widehat{\tau}^{adv}$, in \%)} }   \\
\hline
Mean  & 2.6   & 2.8   & 2.6   & 2.5   & 2.2   & 2.5   & 2.5   & 2.7   & 2.4   & 2.4   & 2.7   & 2.6   & 2.3   & 2.5   & 2.1   & 2.2   & 1.9   & 1.8   & 1.8   & 2.2 \\
Median & 2.2   & 2.5   & 2.2   & 2.2   & 1.9   & 2.1   & 2.1   & 2.6   & 2.3   & 1.9   & 2.8   & 2.2   & 1.9   & 2.3   & 1.8   & 2.0   & 1.8   & 1.6   & 1.4   & 1.8 \\
\hline
\multicolumn{20}{l}{\textit{Additive term ($\widehat{t}^{add}/\widetilde{p}$, in \%)} }   \\
\hline
Mean  & 2.9   & 2.7   & 2.5   & 2.5   & 3.2   & 2.8   & 2.8   & 2.4   & 2.6   & 3.2   & 2.9   & 3.0   & 2.8   & 2.4   & 2.4   & 2.1   & 2.5   & 1.9   & 1.9   & 1.5 \\
Median & 2.0   & 1.8   & 1.6   & 1.3   & 2.0   & 2.0   & 2.2   & 1.6   & 2.0   & 2.5   & 1.9   & 2.2   & 1.9   & 1.8   & 2.1   & 1.7   & 1.9   & 1.6   & 1.6   & 0.8 \\
\hline
\# observations & 31865 & 32146 & 32344 & 33181 & 33986 & 34585 & 36090 & 36407 & 37255 & 37672 & 37757 & 41431 & 41763 & 39604 & 38950 & 37332 & 37748 & 38562 & 38387 & 38473 \\
\hline
\hline
\end{tabular}}

\end{center}
\label{tab_oa:result_vessel_ally3}%
\end{table}
\end{landscape}

\setcounter{table}{0}
\renewcommand{\thetable}{C.\arabic{table}}

\subsection{Eliminating the composition effects: Primary vs. Maufacturing sector} \label{sec_oa:comp-effects}

%\todo{A mon avis, a mettre en online appendix. Expliquer comment se fait la decomposition en secteurs primaire/manuf.}

In this Section, we characterize the time trend in international transport costs at a more disaggregated level, by distinguishing the trade flows for primary goods and manufactured goods. The evolution in transport costs over time, by transport mode (overall transport costs and composition effects excluded) are reported in \ref{fig:totalTC_compeffects_excl_manuf} for the manufacturing sector, and in Figure \ref{fig:totalTC_compeffects_excl_primary} for the primary goods.

\begin{figure}[htbp]
\caption{Transport costs (with and without composition effects), Manufacturing}
\label{fig:totalTC_compeffects_excl_manuf}
\begin{center}
\includegraphics[height=4in]
{graph_composition_manuf.pdf}
\end{center}
\end{figure}

\begin{figure}[htbp]
\caption{Transport costs (with and without composition effects), Primary goods}
\label{fig:totalTC_compeffects_excl_primary}
\begin{center}
\includegraphics[height=4in]
{graph_composition_primary.pdf}
\end{center}
\end{figure}

\begin{figure}[htbp]
\caption{Share of primary goods in the value of total US imports}
\label{fig:Share_prim_goods}
\begin{center}
\includegraphics[height=4in]
{Share_of_primary.pdf}
\end{center}
\end{figure}

%\textbf{A commenter. A mettre en Online Appendix?}

[TO BE COMPLETED]

\end{document}

\subsubsection{4-Digits Level Product Classification}

\begin{landscape}
\begin{table}[htbp]
\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
\caption{Air: Transport costs estimates, all years, 4-digit}
\begin{center}
\scalebox{0.85}{\begin{tabular}{lcccccccccccccccccccc}
\hline
\hline

%%%%%%%%%%%%%%%%2nd part of the table

\hline\hline
\multicolumn{20}{c}{ }  \\
\multicolumn{20}{c}{Continued}  \\
\hline\hline

\hline
\hline
\end{tabular}}

\end{center}
\label{tab_oa:result_vessel_ally3}%
\end{table}
\end{landscape}