### 2018-09-26 redaction/online_appendix_travail.tex
Travail Guillaume Online Appendix

\documentclass[11pt,twoside, authoryear]{elsarticle}
\usepackage[frenchb, english]{babel}
\usepackage[utf8]{inputenc}
%\usepackage[utf8]{fontenc}
%\usepackage[ansinew]{inputenc}
\usepackage{setspace}
\usepackage{babel,varioref}
\usepackage{supertabular}
\usepackage{graphicx}
%\usepackage[pdftex]{graphicx}
%\DeclareGraphicsExtensions{.pdf,.mps,.png,.jpg,.eps}
\usepackage{lscape}
\usepackage[authoryear]{natbib}
%\usepackage[frenchb,english]{babel}
\usepackage{multirow}
\usepackage{ifthen}
%\usepackage{harvard}
\usepackage{vmargin}
\usepackage{verbatim}
\usepackage{array}
%\usepackage[authoryear]{natbib}
\usepackage{hyperref}
\hypersetup{colorlinks=true,linkcolor=Black, citecolor=blue, urlcolor=blue}
%\setmarginsrb{2cm}{2cm}{2cm}{2cm}{0cm}{0cm}{0cm}{0cm}
\usepackage{booktabs,caption,fixltx2e}

\usepackage[flushleft]{threeparttable}

%\usepackage{titling}

%\setlength{\droptitle}{-10em}   % This is your set screw

\makeatletter
\def\ps@pprintTitle{%
  \let\@oddhead\@empty
  \let\@evenhead\@empty
  \let\@oddfoot\@empty
  \let\@evenfoot\@oddfoot
}
\makeatother
\usepackage{adjustbox}
\usepackage{chngcntr}
\counterwithin{table}{section}
%\usepackage[round]{natbib}

\renewcommand{\thesection}{\Alph{section}}

% \newcommand\section[1]{%
%   \refstepcounter{section}%
%   \addcontentsline{toc}{section}{\protect\numberline{\thesection}#1}%
%   \sectionmark{#1}}
%   }

% % \refstepcounter{section}%
% %   \addcontentsline{toc}{section}{\protect\numberline{\thesection}#1}%
% %   \sectionmark{#1}}

%   \newcommand\subsection[1]{%
%   \refstepcounter{subsection}%
%   \addcontentsline{toc}{subsection}{\protect\numberline{\thesubsection}#1}%
%   %\subsectionmark{#1}
%   }

%\EnableSectionsInLOFT

\begin{document}
\title{\textbf{Beyond the Iceberg Hypothesis: \\Opening the Black Box of Transport Costs}\\Online Appendix}

\urlstyle{rm}
\author{Guillaume Daudin\footnote{\noindent Universit\'{e} Paris-Dauphine, PSL Research University, LEDa-DIAL, UMR 225, France \&
SciencesPo, Observatoire Français des Conjonctures \'{E}conomiques (OFCE), France; email: \url{guillaume.daudin@dauphine.fr}} \qquad \quad Jérôme Héricourt\footnote{\noindent Université de Lille - LEM-CNRS (UMR 9221) and CEPII; email: \url{jerome.hericourt@univ-lille1.fr}} \qquad \quad Lise Patureau\footnote{\noindent Corresponding author. Universit\'{e} Paris-Dauphine, PSL Research University, LEDa, France ; email: \url{lise.patureau@dauphine.fr}}}
%\date{\vspace{-5ex}}

%Modèle additif seul DONE
%+Tableaux avec toutes les années en 3 digit DONE
% Tableaux avec toutes les années en 4 digit ???
%Harmoniser la présentation des résultats (cf table 1 du papier) DONE
%Les résultats sur la version Hummels mais avec sa pondération: en annexe du papier? En online appendix? DONE, mais à compléter

\maketitle

%\newpage

\tableofcontents
\vspace{1cm}
\newpage
\listoftables

\newpage

%
%
%
%
%			SECTION A / ADDITIVE ONLY
%			
%			
%				
%
%

	\renewcommand\thesubsubsection{\Alph{subsection}.\arabic{subsubsection}}

	\renewcommand\thesubsection{\Alph{subsection}}

\subsection{Three models: Comparison \label{secoa:additive_only}}

The paper compares the empirical performances of two models: one with only ad-valorem costs (Model (A)) and one with where both ad-valorem and additive costs (Model (B)).
In the interest of comprehensiveness, this appendix looks at a model with only additive costs (Model (C)).
The estimated equation is: $\ln\left(\frac{p_{ik}}{\widetilde{p}_{ik}}-1 \right)= \ln \left(\frac{t_{i} + t_{s(k)}}{\widetilde{p}_{ik}}\right) + \epsilon^{add}_{ik}$. \\
We first compare estimates of the transport costs components under the three models. In a second step, we report quality of fit tests for the three models.

\subsubsection{Estimation results}

In this section, we report the estimation results of the three models. We also report the estimation results for Model (C), where only additive costs are modeled. Table \ref{tab:3models_estimation_results_air} reports the results for air transport, Table \ref{tab:3models_estimation_results_vessel} for vessel transport.

%%%% RAW TABLE TO BE FOUND AT: "...\Dropbox\Papier_Lise_Guillaume\trade_cost\results\3_models"
% modele iceberg tout seul : terme nlI
% colonnes L à U: modèle avec les deux
% termes nla, de V à Z, modèles avec que de l'additif
%nlI quer iceberg
%nla que additif
%nl : les deux
\setcounter{table}{0}
\renewcommand{\thetable}{A.\arabic{table}}

\begin{table}[htbp]
  \centering
   \footnotesize{
    \caption{Estimation results of the three models (Air, 3-digit level)}
  \label{tab:3models_estimation_results_air}%
    \begin{tabular}{l|c c c c c c c c c}
    \hline \hline
    \textbf{Year} & \multicolumn{1}{c}{\textbf{1974}} & \multicolumn{1}{c}{\textbf{1980}} & \multicolumn{1}{c}{\textbf{1985}} & \multicolumn{1}{c}{\textbf{1990}} & \multicolumn{1}{c}{\textbf{1995}} & \multicolumn{1}{c}{\textbf{2000}} & \multicolumn{1}{c}{\textbf{2005}} & \multicolumn{1}{c}{\textbf{2010}} & \multicolumn{1}{c}{\textbf{2013}} \\ \hline
   $\#$ obs. & \multicolumn{1}{c}{14955} & \multicolumn{1}{c}{16118} & \multicolumn{1}{c}{19908} & \multicolumn{1}{c}{24958} & \multicolumn{1}{c}{31037} & \multicolumn{1}{c}{35027} & \multicolumn{1}{c}{41806} & \multicolumn{1}{c}{40279} & \multicolumn{1}{c}{39351} \\
   $\#$  origin countries & \multicolumn{1}{c}{152} & \multicolumn{1}{c}{165} & \multicolumn{1}{c}{169} & \multicolumn{1}{c}{181} & \multicolumn{1}{c}{207} & \multicolumn{1}{c}{208} & \multicolumn{1}{c}{211} & \multicolumn{1}{c}{210} & \multicolumn{1}{c}{210} \\
   $\#$  sectors & \multicolumn{1}{c}{203} & \multicolumn{1}{c}{204} & \multicolumn{1}{c}{207} & \multicolumn{1}{c}{212} & \multicolumn{1}{c}{217} & \multicolumn{1}{c}{218} & \multicolumn{1}{c}{217} & \multicolumn{1}{c}{216} & \multicolumn{1}{c}{212} \\ \hline
    \multicolumn{10}{l}{\textbf{Model (A) - Iceberg transport costs only} ($\widehat{\tau}^{ice}$)} \\ \hline
    Mean (in \%) & \multicolumn{1}{c}{6.93} & \multicolumn{1}{c}{5.41} & \multicolumn{1}{c}{6.08} & \multicolumn{1}{c}{5.03} & \multicolumn{1}{c}{4.61} & \multicolumn{1}{c}{3.60} & \multicolumn{1}{c}{4.10} & \multicolumn{1}{c}{4.19} & \multicolumn{1}{c}{3.36} \\
    Median (in \%)& \multicolumn{1}{c}{5.43} & \multicolumn{1}{c}{3.79} & \multicolumn{1}{c}{5.47} & \multicolumn{1}{c}{4.37} & \multicolumn{1}{c}{3.80} & \multicolumn{1}{c}{2.47} & \multicolumn{1}{c}{3.12} & \multicolumn{1}{c}{3.41} & \multicolumn{1}{c}{2.92} \\
    Std & \multicolumn{1}{c}{0.04} & \multicolumn{1}{c}{0.04} & \multicolumn{1}{c}{0.04} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} \\ \hline
    \multicolumn{10}{l}{\textbf{Model (B) - With additive and ad-valorem transport costs}} \\ \hline
     \multicolumn{10}{l}{ \textit{Multiplicative term} ($\widehat{\tau}^{adv}$)} \\ \hline
    Mean (in \%) & \multicolumn{1}{c}{3.64} & \multicolumn{1}{c}{2.32} & \multicolumn{1}{c}{2.46} & \multicolumn{1}{c}{2.38} & \multicolumn{1}{c}{2.05} & \multicolumn{1}{c}{1.66} & \multicolumn{1}{c}{2.00} & \multicolumn{1}{c}{2.57} & \multicolumn{1}{c}{1.70} \\
    Median (in \%)& \multicolumn{1}{c}{2.71} & \multicolumn{1}{c}{1.57} & \multicolumn{1}{c}{1.79} & \multicolumn{1}{c}{1.60} & \multicolumn{1}{c}{1.39} & \multicolumn{1}{c}{1.20} & \multicolumn{1}{c}{1.57} & \multicolumn{1}{c}{2.24} & \multicolumn{1}{c}{1.72} \\
    Std & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.01} \\ \hline
      \multicolumn{10}{l}{\textit{Additive term} ($\widehat{t}/\widetilde{p}$)} \\ \hline
    Mean (in \%) & \multicolumn{1}{c}{2.56} & \multicolumn{1}{c}{2.04} & \multicolumn{1}{c}{2.83} & \multicolumn{1}{c}{1.83} & \multicolumn{1}{c}{1.64} & \multicolumn{1}{c}{1.30} & \multicolumn{1}{c}{1.43} & \multicolumn{1}{c}{1.13} & \multicolumn{1}{c}{1.01} \\
    Median (in \%)& \multicolumn{1}{c}{1.13} & \multicolumn{1}{c}{0.54} & \multicolumn{1}{c}{1.30} & \multicolumn{1}{c}{0.84} & \multicolumn{1}{c}{0.68} & \multicolumn{1}{c}{0.45} & \multicolumn{1}{c}{0.53} & \multicolumn{1}{c}{0.43} & \multicolumn{1}{c}{0.47} \\
    Std & \multicolumn{1}{c}{0.04} & \multicolumn{1}{c}{0.04} & \multicolumn{1}{c}{0.04} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.03} & \multicolumn{1}{c}{0.02} & \multicolumn{1}{c}{0.02} \\ \hline
    \multicolumn{10}{l}{\textbf{Model (C) - With additive transport costs only} ($\widehat{t}^{add}/\widetilde{p}$)} \\ \hline
    Mean (in \%) & \multicolumn{1}{c}{6.88} & \multicolumn{1}{c}{4.82} & \multicolumn{1}{c}{5.99} & \multicolumn{1}{c}{4.44} & \multicolumn{1}{c}{3.80} & \multicolumn{1}{c}{3.09} & \multicolumn{1}{c}{4.00} & \multicolumn{1}{c}{4.49} & \multicolumn{1}{c}{3.29} \\
    Median (in \%)& \multicolumn{1}{c}{4.45} & \multicolumn{1}{c}{1.82} & \multicolumn{1}{c}{3.36} & \multicolumn{1}{c}{2.27} & \multicolumn{1}{c}{1.72} & \multicolumn{1}{c}{1.39} & \multicolumn{1}{c}{1.93} & \multicolumn{1}{c}{2.72} & \multicolumn{1}{c}{2.06} \\
    Std & \multicolumn{1}{c}{0.09} & \multicolumn{1}{c}{0.08} & \multicolumn{1}{c}{0.08} & \multicolumn{1}{c}{0.10} & \multicolumn{1}{c}{0.09} & \multicolumn{1}{c}{0.06} & \multicolumn{1}{c}{0.07} & \multicolumn{1}{c}{0.07} & \multicolumn{1}{c}{0.05} \\  \hline \hline
    \end{tabular}
    }
\end{table}%

\begin{table}[htbp]
  \centering
   \footnotesize{
    \caption{Estimation results of the three models (Vessel, 3-digit level)}
  \label{tab:3models_estimation_results_vessel}%
    \begin{tabular}{l|c c c c c c c c c}
    \hline \hline
    \textbf{Year} & 1974  & 1980  & 1985  & 1990  & 1995  & 2000  & 2005  & 2010  & 2013 \\ \hline
    $\# $obs. & 19007 & 17356 & 23348 & 28383 & 32146 & 36090 & 41319 & 37748 & 38473 \\
    $\#$  origin countries & 154   & 163   & 171   & 179   & 201   & 206   & 206   & 198   & 203 \\
    $\#$  sectors & 239   & 232   & 232   & 232   & 228   & 230   & 231   & 226   & 224 \\ \hline
    \multicolumn{10}{l}{\textbf{Model (A) - Iceberg transport costs only} ($\widehat{\tau}^{ice}$)} \\ \hline
    Mean (in \%) & 9.79  & 6.53  & 6.88  & 5.67  & 5.14  & 5.10  & 5.47  & 3.99  & 3.60 \\
    Median (in \%)& 9.58  & 5.50  & 6.33  & 4.63  & 4.29  & 4.85  & 4.90  & 3.56  & 3.28 \\
    Std & 0.05  & 0.04  & 0.04  & 0.03  & 0.03  & 0.03  & 0.03  & 0.02  & 0.02 \\ \hline
    \multicolumn{10}{l}{\textbf{Model (B ) - With additive and ad-valorem transport costs} } \\ \hline
     \multicolumn{10}{l}{\textit{Multiplicative term} ($\widehat{\tau}^{adv}$)} \\ \hline
    Mean (in \%) & 5.42  & 3.08  & 4.02  & 3.31  & 2.79  & 2.49  & 2.68  & 1.95  & 2.22 \\
    Median (in \%) & 4.93  & 2.42  & 3.60  & 2.81  & 2.53  & 2.07  & 2.08  & 1.76  & 1.82 \\
    Std & 0.04  & 0.02  & 0.03  & 0.02  & 0.02  & 0.02  & 0.02  & 0.02  & 0.01 \\ \hline
       \multicolumn{10}{l}{\textit{Additive term} ($\widehat{t}/\widetilde{p}$)}  \\ \hline
    Mean (in \%)  & 5.08  & 3.38  & 3.19  & 2.73  & 2.73  & 2.80  & 3.02  & 2.47  & 1.46 \\
    Median (in \%) & 2.94  & 2.27  & 2.06  & 1.70  & 1.82  & 2.19  & 2.16  & 1.89  & 0.76 \\
    Std & 0.09  & 0.05  & 0.04  & 0.04  & 0.04  & 0.04  & 0.04  & 0.03  & 0.02 \\ \hline
    \multicolumn{10}{l}{\textbf{Model (C) - With additive transport costs only} ($\widehat{t}^{add}/\widetilde{p}$)} \\ \hline
    Mean (in \%) & 14.51 & 10.05 & 10.47 & 14.62 & 8.37  & 8.02  & 8.41  & 6.40  & 5.23 \\
    Median (in \%)& 9.53  & 6.69  & 7.16  & 6.22  & 4.68  & 4.93  & 5.74  & 3.87  & 3.57 \\
    Std & 0.24  & 0.17  & 0.18  & 0.30  & 0.15  & 0.16  & 0.15  & 0.15  & 0.10 \\ \hline \hline
   \end{tabular}%
   }
\end{table}%

Results for Models (A) (ad-valorem costs only) and (B) (both ad-valorem and additive costs) are identical to those reported in the paper. Somewhat unsurprisingly, the estimated size of transport costs under Model (C) is of same order of magnitude as in Models (A) and (B). We also observe a downward trend of transport costs over time, in particular since 1980.\\

\subsubsection{Quality of fit diagnostic tests}

To go further into characterizing the empirical relevance of the model with aditive costs only (Model (C)), we compare quality of fit diagnostic tests in this section. Table \ref{tab:3models_diagnosis_air} reports the results for air transport, and Table \ref{tab:3models_diagnosis_vessel} those for vessel transport. In both tables, we report the values of the $R^2$, the Standard Error of Regression (SER), the AIC criterion  and the log-likelihood (LL) value. We also report the value of the log-likelihood ratio that tests the quality of fit of the global model (Model (B)) compared to the other two models.

\begin{landscape}
\begin{table}[htbp]
  \centering
  \footnotesize{
  \caption{Quality-of-fit diagnostic tests of the three models (Air, 3-digit level)}
    \label{tab:3models_diagnosis_air}%
   \begin{tabular}{l|c|c|c|c|c|c|c|c|c}
    \hline \hline
    \textbf{Year} & \multicolumn{1}{c}{\textbf{1974}} & \multicolumn{1}{c}{\textbf{1980}} & \multicolumn{1}{c}{\textbf{1985}} & \multicolumn{1}{c}{\textbf{1990}} & \multicolumn{1}{c}{\textbf{1995}} & \multicolumn{1}{c}{\textbf{2000}} & \multicolumn{1}{c}{\textbf{2005}} & \multicolumn{1}{c}{\textbf{2010}} & \multicolumn{1}{c}{\textbf{2013}} \\ \hline
    \multicolumn{10}{l}{\textbf{$R^2$}} \\ \hline
    Model (A) & \multicolumn{1}{c}{0.297} & \multicolumn{1}{c}{0.267} & \multicolumn{1}{c}{0.302} & \multicolumn{1}{c}{0.251} & \multicolumn{1}{c}{0.142} & \multicolumn{1}{c}{0.318} & \multicolumn{1}{c}{0.460} & \multicolumn{1}{c}{0.421} & \multicolumn{1}{c}{0.313} \\
    Model (B) & \multicolumn{1}{c}{0.594} & \multicolumn{1}{c}{0.646} & \multicolumn{1}{c}{0.635} & \multicolumn{1}{c}{0.627} & \multicolumn{1}{c}{0.658} & \multicolumn{1}{c}{0.640} & \multicolumn{1}{c}{0.593} & \multicolumn{1}{c}{0.513} & \multicolumn{1}{c}{0.419} \\
    Model (C) & \multicolumn{1}{c}{0.489} & \multicolumn{1}{c}{0.543} & \multicolumn{1}{c}{0.531} & \multicolumn{1}{c}{0.517} & \multicolumn{1}{c}{0.546} & \multicolumn{1}{c}{0.518} & \multicolumn{1}{c}{0.464} & \multicolumn{1}{c}{0.339} & \multicolumn{1}{c}{0.295} \\ \hline
    \multicolumn{10}{l}{\textbf{SER}}\\ \hline
    Model (A) & \multicolumn{1}{c}{0.791} & \multicolumn{1}{c}{0.860} & \multicolumn{1}{c}{0.831} & \multicolumn{1}{c}{0.811} & \multicolumn{1}{c}{0.798} & \multicolumn{1}{c}{0.844} & \multicolumn{1}{c}{0.837} & \multicolumn{1}{c}{0.857} & \multicolumn{1}{c}{0.920} \\
    Model (B) & \multicolumn{1}{c}{0.674} & \multicolumn{1}{c}{0.715} & \multicolumn{1}{c}{0.692} & \multicolumn{1}{c}{0.675} & \multicolumn{1}{c}{0.641} & \multicolumn{1}{c}{0.697} & \multicolumn{1}{c}{0.727} & \multicolumn{1}{c}{0.787} & \multicolumn{1}{c}{0.847} \\
    Model (C) & \multicolumn{1}{c}{1.610} & \multicolumn{1}{c}{1.778} & \multicolumn{1}{c}{1.736} & \multicolumn{1}{c}{1.699} & \multicolumn{1}{c}{1.700} & \multicolumn{1}{c}{1.786} & \multicolumn{1}{c}{1.783} & \multicolumn{1}{c}{1.776} & \multicolumn{1}{c}{1.723} \\ \hline
   \multicolumn{10}{l}{ \textbf{AIC criteria}} \\ \hline
    Model (A) & \multicolumn{1}{c}{35675.0} & \multicolumn{1}{c}{41171.0} & \multicolumn{1}{c}{49315.0} & \multicolumn{1}{c}{60715.6} & \multicolumn{1}{c}{74386.4} & \multicolumn{1}{c}{87492.5} & \multicolumn{1}{c}{103983.0} & \multicolumn{1}{c}{102297.7} & \multicolumn{1}{c}{106130.6} \\
    Model (B ) & \multicolumn{1}{c}{31387.3} & \multicolumn{1}{c}{35738.4} & \multicolumn{1}{c}{42535.8} & \multicolumn{1}{c}{52098.9} & \multicolumn{1}{c}{61343.7} & \multicolumn{1}{c}{74954.9} & \multicolumn{1}{c}{92758.6} & \multicolumn{1}{c}{95887.1} & \multicolumn{1}{c}{100155.4} \\
    Model (C) & \multicolumn{1}{c}{40808.1} & \multicolumn{1}{c}{45138.5} & \multicolumn{1}{c}{55214.8} & \multicolumn{1}{c}{69458.5} & \multicolumn{1}{c}{83958.6} & \multicolumn{1}{c}{100040.8} & \multicolumn{1}{c}{123592.1} & \multicolumn{1}{c}{129359.0} & \multicolumn{1}{c}{127399.2} \\ \hline
    \multicolumn{10}{l}{\textbf{Log-likelihood}}\\ \hline
    Model (A) & \multicolumn{1}{c}{-17530.5} & \multicolumn{1}{c}{-20253.5} & \multicolumn{1}{c}{-24315.5} & \multicolumn{1}{c}{-29977.8} & \multicolumn{1}{c}{-36811.2} & \multicolumn{1}{c}{-43341.3} & \multicolumn{1}{c}{-51648.5} & \multicolumn{1}{c}{-50746.8} & \multicolumn{1}{c}{-52690.3} \\
    Model (B ) & \multicolumn{1}{c}{-15125.6} & \multicolumn{1}{c}{-17263.2} & \multicolumn{1}{c}{-20686.9} & \multicolumn{1}{c}{-25393.5} & \multicolumn{1}{c}{-30036.9} & \multicolumn{1}{c}{-36788.4} & \multicolumn{1}{c}{-45768.3} & \multicolumn{1}{c}{-47277.5} & \multicolumn{1}{c}{-49419.7} \\
    Model (C) & \multicolumn{1}{c}{-20074.1} & \multicolumn{1}{c}{-22217.2} & \multicolumn{1}{c}{-27251.4} & \multicolumn{1}{c}{-34355.3} & \multicolumn{1}{c}{-41634.3} & \multicolumn{1}{c}{-49625.4} & \multicolumn{1}{c}{-61533.0} & \multicolumn{1}{c}{-64339.5} & \multicolumn{1}{c}{-63316.6} \\ \hline
    \multicolumn{10}{l}{\textbf{Test LL} }\\ \hline
    Stat LL ratio (B vs A) & \multicolumn{1}{c}{4809.7} & \multicolumn{1}{c}{5980.6} & \multicolumn{1}{c}{7257.3} & \multicolumn{1}{c}{9168.7} & \multicolumn{1}{c}{13548.7} & \multicolumn{1}{c}{13105.7} & \multicolumn{1}{c}{11760.4} & \multicolumn{1}{c}{6938.6} & \multicolumn{1}{c}{6541.2} \\
    \textbackslash{}\# restrictions & \multicolumn{1}{c}{16929} & \multicolumn{1}{c}{18098} & \multicolumn{1}{c}{21893} & \multicolumn{1}{c}{26948} & \multicolumn{1}{c}{33032} & \multicolumn{1}{c}{37027} & \multicolumn{1}{c}{43811} & \multicolumn{1}{c}{42289} & \multicolumn{1}{c}{41364} \\
    p-value & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} \\ \hline
    Stat LL ratio (B vs C) & \multicolumn{1}{c}{4948.4} & \multicolumn{1}{c}{4954.0} & \multicolumn{1}{c}{6564.5} & \multicolumn{1}{c}{8961.8} & \multicolumn{1}{c}{11597.4} & \multicolumn{1}{c}{12837.0} & \multicolumn{1}{c}{15764.7} & \multicolumn{1}{c}{17062.0} & \multicolumn{1}{c}{13896.9} \\
    \textbackslash{}\# restrictions & \multicolumn{1}{c}{16929} & \multicolumn{1}{c}{18098} & \multicolumn{1}{c}{21893} & \multicolumn{1}{c}{26948} & \multicolumn{1}{c}{33032} & \multicolumn{1}{c}{37027} & \multicolumn{1}{c}{43811} & \multicolumn{1}{c}{42289} & \multicolumn{1}{c}{41364} \\
    p-value & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} & \multicolumn{1}{c}{0.00} \\ \hline \hline
\end{tabular}%
}
\end{table}%
\end{landscape}

\begin{landscape}
\begin{table}[htbp]
  \centering
  \footnotesize{
  \caption{Quality-of-fit diagnostic tests of the three models (Vessel, 3-digit level)}
    \label{tab:3models_diagnosis_vessel}%
   \begin{tabular}{l|c c c c c c c c c}
    \hline \hline
    \textbf{Year} & \multicolumn{1}{c}{\textbf{1974}} & \multicolumn{1}{c}{\textbf{1980}} & \multicolumn{1}{c}{\textbf{1985}} & \multicolumn{1}{c}{\textbf{1990}} & \multicolumn{1}{c}{\textbf{1995}} & \multicolumn{1}{c}{\textbf{2000}} & \multicolumn{1}{c}{\textbf{2005}} & \multicolumn{1}{c}{\textbf{2010}} & \multicolumn{1}{c}{\textbf{2013}} \\ \hline
    \multicolumn{10}{l}{\textbf{$R^2$}} \\ \hline
    Model (A) & 0.450 & 0.415 & 0.427 & 0.456 & 0.438 & 0.401 & 0.378 & 0.350 & 0.339 \\
    Model (B) & 0.612 & 0.575 & 0.571 & 0.590 & 0.611 & 0.571 & 0.541 & 0.491 & 0.462 \\
    Model (C ) & 0.424 & 0.401 & 0.374 & 0.429 & 0.456 & 0.431 & 0.417 & 0.358 & 0.349 \\ \hline
    \multicolumn{10}{l}{\textbf{SER} } \\ \hline
    Model (A) & 0.576 & 0.620 & 0.569 & 0.592 & 0.615 & 0.652 & 0.673 & 0.740 & 0.758 \\
    Model (B) & 0.484 & 0.528 & 0.493 & 0.514 & 0.512 & 0.551 & 0.578 & 0.656 & 0.684 \\
    Model (C ) & 1.271 & 1.339 & 1.283 & 1.326 & 1.302 & 1.319 & 1.336 & 1.392 & 1.410 \\ \hline
    \multicolumn{10}{l}{\textbf{AIC criteria} } \\ \hline
    Model (A) & 33328.81 & 33010.27 & 40275.70 & 51142.62 & 60414.92 & 71365.89 & 85051.02 & 84789.89 & 88191.87 \\
    Model (B) & 27331.52 & 28067.31 & 34170.52 & 43664.74 & 49275.33 & 60475.91 & 73020.09 & 76161.33 & 80873.72 \\
    Model (C ) & 46082.40 & 44370.26 & 58829.71 & 71461.52 & 77052.41 & 88746.51 & 103310.93 & 101166.91 & 104290.27 \\ \hline
    \multicolumn{10}{l}{\textbf{Log-likelihood}}\\ \hline
    Model (A) & -16287.40 & -16129.13 & -19767.85 & -25169.31 & -29790.46 & -35263.95 & -42122.51 & -41998.95 & -43692.93 \\
    Model (B) & -12985.76 & -13353.65 & -16398.26 & -21171.37 & -23905.66 & -29490.96 & -35844.04 & -37418.66 & -39751.86 \\
    Model (C ) & -22674.20 & -21814.13 & -29045.86 & -35403.76 & -38125.20 & -43963.25 & -51245.46 & -50348.45 & -51783.14 \\ \hline
   \multicolumn{10}{l}{ \textbf{Test LL} } \\ \hline
    Stat LL ratio (B vs A) & 6603.28 & 5550.96 & 6739.18 & 7995.88 & 11769.59 & 11545.98 & 12556.94 & 9160.56 & 7882.15 \\
    $\#$ restrictions & 393   & 395   & 403   & 411   & 429   & 436   & 437   & 424   & 427 \\
    p-value & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00 \\ \hline
    Stat LL ratio (B vs C) & 19376,88 & 16920,95 & 25295,20 & 28464,79 & 28439,08 & 28944,59 & 30802,84 & 25859,58 & 24062,55 \\
    $\#$ restrictions & 393   & 395   & 403   & 411   & 429   & 436   & 437   & 424   & 427 \\
    p-value & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00  & 0.00 \\
\hline \hline
    \end{tabular}%
}
\end{table}%
\end{landscape}

The model with additive costs only (Model (C)) is consistently dominated (in terms of quality of fit properties) by the model with multiplicative costs only (Model (A)), which is itself consistently dominated by the complete model (Model (B)), whatever the type of diagnostic test considered. That justify our choice to disregard Model (C) in the main text.

%\subsection{Detailed Results \label{secoa:detailed}}
\newpage
\setcounter{table}{0}
\renewcommand{\thetable}{B.\arabic{table}}

\subsection{Transport Cost Estimates: Yearly Detailed Results}
In this section, we complement Table 1 of the main text by reporting the year-to-year results of the estimation driven at the 3-digit classification level. Table \ref{tab_oa:result_air_ally3} reports the results for each year over 1974-2013 for air transport; Table \ref{tab_oa:result_vessel_ally3} reports similar results for vessel. In both cases, we report the estimated values of the transport costs (weighed mean and median) when only ad-valorem costs are modeled (Model (A)) and when both additive and ad-valorem costs are modeled (Model (B)).

\begin{landscape}
\begin{table}[htbp]
\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
  %\centering
\caption{Air: Transport costs estimates, all years, 3-digit}
\begin{center}
\scalebox{0.85}{\begin{tabular}{lcccccccccccccccccccc}
\hline
\hline
Year  & 1974  & 1975  & 1976  & 1977  & 1978  & 1979  & 1980  & 1981  & 1982  & 1983  & 1984  & 1985  & 1986  & 1987  & 1988  & 1989  & 1990  & 1991  & 1992  & 1993 \\
\hline
\multicolumn{20}{l}{\textbf{Model (A) - With only Ad-Valorem Trade Costs} ($\widehat{\tau}^{ice}$, in \%)} \\
\hline
Mean  & 6.9   & 7.5   & 7.2   & 7.7   & 6.9   & 6.1   & 5.4   & 6.0   & 6.4   & 6.9   & 7.2   & 6.1   & 6.4   & 6.6   & 5.7   & 5.3   & 5.0   & 5.1   & 4.9   & 5.1 \\
Median & 5.4   & 6.4   & 6.9   & 7.1   & 6.3   & 5.3   & 3.8   & 4.9   & 5.3   & 6.1   & 6.7   & 5.5   & 5.9   & 6.3   & 5.3   & 4.6   & 4.4   & 4.5   & 4.5   & 4.4 \\
\hline
\multicolumn{20}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem Trade Costs} }\\ \hline
\multicolumn{20}{l}{\textit{Ad-valorem term ($\widehat{\tau}^{adv}$, in \%)} }   \\
\hline
Mean  & 3.6   & 3.7   & 3.9   & 3.8   & 3.2   & 3.0   & 2.3   & 2.8   & 2.8   & 2.6   & 3.3   & 2.5   & 3.2   & 2.6   & 3.1   & 3.1   & 2.4   & 2.7   & 2.2   & 2.4 \\
Median & 2.7   & 2.7   & 2.9   & 2.7   & 2.1   & 2.4   & 1.6   & 1.8   & 1.9   & 1.9   & 2.7   & 1.8   & 2.1   & 2.0   & 2.0   & 1.9   & 1.6   & 1.5   & 1.5   & 1.6 \\

\hline
\multicolumn{20}{l}{\textit{Additive term ($\widehat{t}^{add}/\widetilde{p}$, in \%)} }   \\
\hline
Mean  & 2.6   & 3.0   & 2.3   & 3.1   & 2.6   & 2.1   & 2.0   & 2.0   & 2.3   & 2.8   & 2.5   & 2.8   & 2.6   & 2.9   & 1.7   & 4.6   & 1.8   & 1.8   & 1.9   & 1.9 \\
Median & 1.1   & 1.2   & 0.9   & 1.3   & 1.1   & 0.7   & 0.5   & 0.6   & 0.8   & 1.0   & 1.0   & 1.3   & 1.3   & 1.5   & 1.0   & 0.7   & 0.8   & 0.6   & 0.9   & 0.8 \\
\hline
\# observations & 14955 & 15299 & 11397 & 10707 & 15222 & 15684 & 16118 & 16864 & 17322 & 18180 & 20644 & 19908 & 20695 & 20793 & 24663 & 25197 & 24958 & 25156 & 26191 & 28296 \\

%%%%%%%%%%%%%%%%2nd part of the table

\hline\hline
\multicolumn{20}{c}{ }  \\
\multicolumn{20}{c}{Continued}  \\
\hline\hline

Year  & 1994  & 1995  & 1996  & 1997  & 1998  & 1999  & 2000  & 2001  & 2002  & 2003  & 2004  & 2005  & 2006  & 2007  & 2008  & 2009  & 2010  & 2011  & 2012  & 2013 \\
\hline
\multicolumn{20}{l}{\textbf{Model (A) - With only Ad-Valorem Trade Costs} ($\widehat{\tau}^{ice}$, in \%)} \\
\hline
Mean  & 4.6   & 4.6   & 4.2   & 4.1   & 3.8   & 3.8   & 3.6   & 3.5   & 3.8   & 3.9   & 4.0   & 4.1   & 3.9   & 4.1   & 4.1   & 4.0   & 4.2   & 3.9   & 3.7   & 3.4 \\
Median & 3.7   & 3.8   & 3.1   & 3.0   & 2.7   & 2.8   & 2.5   & 2.4   & 2.7   & 2.6   & 2.9   & 3.1   & 2.7   & 3.0   & 3.2   & 3.0   & 3.4   & 3.1   & 3.0   & 2.9 \\
\hline
\multicolumn{20}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem Trade Costs} }\\ \hline
\multicolumn{20}{l}{\textit{Ad-valorem term ($\widehat{\tau}^{adv}$, in \%)} }   \\
\hline
Mean  & 2.3   & 2.1   & 1.9   & 1.8   & 1.8   & 1.8   & 1.7   & 1.6   & 1.6   & 1.9   & 1.9   & 2.0   & 1.8   & 2.3   & 2.3   & 2.3   & 2.6   & 2.2   & 2.2   & 1.7 \\
Median & 1.3   & 1.4   & 1.4   & 1.3   & 1.3   & 1.5   & 1.2   & 1.1   & 1.2   & 1.4   & 1.4   & 1.6   & 1.4   & 1.9   & 1.9   & 1.8   & 2.2   & 1.7   & 1.9   & 1.7 \\
\hline
\multicolumn{20}{l}{\textit{Additive term ($\widehat{t}^{add}/\widetilde{p}$, in \%)} }   \\
\hline
Mean  & 1.7   & 1.6   & 1.5   & 1.5   & 1.4   & 1.4   & 1.3   & 1.3   & 1.6   & 1.4   & 1.5   & 1.4   & 1.3   & 1.2   & 1.2   & 1.2   & 1.1   & 1.1   & 0.9   & 1.0 \\
Median & 0.8   & 0.7   & 0.6   & 0.6   & 0.5   & 0.5   & 0.5   & 0.5   & 0.5   & 0.5   & 0.6   & 0.5   & 0.5   & 0.5   & 0.5   & 0.5   & 0.4   & 0.4   & 0.4   & 0.5 \\
\hline
\# observations & 29948 & 31037 & 32187 & 33502 & 33492 & 33523 & 35027 & 34885 & 35159 & 35891 & 36990 & 41806 & 42554 & 40858 & 40159 & 38275 & 40279 & 41190 & 40909 & 39351 \\

\hline
\hline
\end{tabular}}%

\end{center}
\label{tab_oa:result_air_ally3}%
\end{table}%
\end{landscape}

As mentioned in the paper, the estimates for air transport costs in 1989 show a surprisingly high value for the additive component (the additive cost is estimated to amount to $4.6\%$ of the export price, whereas it amounts to 2.5\% on average between 1974 and 1988, and to 1.7\% over the following decade 1990-2000). This can be attributed to the presence of outliers in the distribution of the additive costs estimates. The maximum value for $\widehat{t}/\widetilde{p}$ is 10,000\% in 1989, whereas it amounts to 1,690\% on average over 1974-1988 and to 1,500\% on average over 1990-2000. Accordingly, in the paper we discard this year 1989 when we report the average values over the period of the transport costs estimates in air transport.\\
Still, this does not make much difference. When 1989 is included, the weighed mean transport cost value amounts to 1.9\% for the ad-valorem component, vs 1.8\% when 1989 is excluded. The weighed median value is left unchanged at 2.9\% whether 1989 is included or not.

\begin{landscape}
\begin{table}[htbp]
\def\sym#1{\ifmmode^{#1}\else\(^{#1}\)\fi}
\caption{Vessel: Transport costs estimates, all years, 3-digit}
\begin{center}
\scalebox{0.85}{\begin{tabular}{lcccccccccccccccccccc}
\hline
\hline
Year  & 1974  & 1975  & 1976  & 1977  & 1978  & 1979  & 1980  & 1981  & 1982  & 1983  & 1984  & 1985  & 1986  & 1987  & 1988  & 1989  & 1990  & 1991  & 1992  & 1993 \\
\hline
\multicolumn{20}{l}{\textbf{Model (A) - With only Ad-Valorem Trade Costs} ($\widehat{\tau}^{ice}$, in \%)} \\
\hline
Mean  & 9.8   & 9.9   & 8.9   & 8.3   & 8.1   & 7.5   & 6.5   & 6.0   & 6.3   & 7.0   & 7.0   & 6.9   & 6.7   & 6.2   & 6.1   & 5.7   & 5.7   & 5.5   & 5.0   & 5.2 \\
Median & 9.6   & 8.5   & 8.0   & 7.3   & 7.1   & 6.5   & 5.5   & 5.0   & 5.9   & 5.7   & 6.1   & 6.3   & 7.0   & 6.3   & 5.7   & 4.8   & 4.6   & 4.4   & 4.2   & 4.7 \\
\hline
\multicolumn{20}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem Trade Costs} }\\ \hline
\multicolumn{20}{l}{\textit{Ad-valorem term ($\widehat{\tau}^{adv}$, in \%)} }   \\
\hline
Mean  & 5.4   & 4.8   & 5.4   & 5.2   & 5.9   & 4.6   & 3.1   & 3.3   & 3.4   & 4.2   & 4.1   & 4.0   & 3.9   & 3.6   & 4.0   & 3.0   & 3.3   & 3.0   & 2.6   & 2.9 \\
Median & 4.9   & 4.1   & 4.8   & 4.4   & 5.4   & 4.0   & 2.4   & 2.9   & 2.9   & 3.9   & 3.5   & 3.6   & 3.6   & 3.0   & 3.5   & 2.6   & 2.8   & 2.7   & 2.3   & 2.6 \\
\multicolumn{20}{l}{\textit{Additive term ($\widehat{t}^{add}/\widetilde{p}$, in \%)} }   \\
\hline
Mean  & 5.1   & 5.5   & 3.5   & 3.5   & 2.5   & 3.1   & 3.4   & 2.9   & 3.5   & 2.9   & 3.2   & 3.2   & 2.9   & 2.8   & 2.4   & 2.9   & 2.7   & 2.8   & 2.7   & 2.7 \\
Median & 2.9   & 3.6   & 1.9   & 1.7   & 1.2   & 1.7   & 2.3   & 1.5   & 2.3   & 2.0   & 2.3   & 2.1   & 1.8   & 1.8   & 1.3   & 2.0   & 1.7   & 1.7   & 1.8   & 1.6 \\
\hline
\# observations & 19007 & 18710 & 13615 & 12826 & 16601 & 17274 & 17356 & 17788 & 18075 & 18883 & 21650 & 23348 & 23729 & 23626 & 27661 & 29106 & 28383 & 28095 & 29050 & 30839 \\ %%%%%%%%%%%%%%%%2nd part of the table

\hline\hline
\multicolumn{20}{c}{ }  \\
\multicolumn{20}{c}{Continued}  \\
\hline\hline

Year  & 1994  & 1995  & 1996  & 1997  & 1998  & 1999  & 2000  & 2001  & 2002  & 2003  & 2004  & 2005  & 2006  & 2007  & 2008  & 2009  & 2010  & 2011  & 2012  & 2013 \\
\hline
\multicolumn{20}{l}{\textbf{Model (A) - With only Ad-Valorem Trade Costs} ($\widehat{\tau}^{ice}$, in \%)} \\
\hline
Mean  & 5.2   & 5.1   & 4.8   & 4.7   & 4.8   & 5.0   & 5.1   & 5.0   & 4.8   & 5.2   & 5.4   & 5.5   & 4.8   & 4.7   & 4.4   & 4.3   & 4.0   & 3.5   & 3.6   & 3.6 \\
Median & 4.1   & 4.3   & 3.9   & 3.9   & 3.9   & 4.5   & 4.9   & 4.6   & 4.1   & 4.8   & 5.1   & 4.9   & 4.2   & 4.2   & 3.8   & 4.1   & 3.6   & 3.0   & 3.1   & 3.3 \\
\hline
\multicolumn{20}{l}{\textbf{Model (B) - With Additive \& Ad-Valorem Trade Costs} }\\ \hline
\multicolumn{20}{l}{\textit{Ad-valorem term ($\widehat{\tau}^{adv}$, in \%)} }   \\
\hline
Mean  & 2.6   & 2.8   & 2.6   & 2.5   & 2.2   & 2.5   & 2.5   & 2.7   & 2.4   & 2.4   & 2.7   & 2.6   & 2.3   & 2.5   & 2.1   & 2.2   & 1.9   & 1.8   & 1.8   & 2.2 \\
Median & 2.2   & 2.5   & 2.2   & 2.2   & 1.9   & 2.1   & 2.1   & 2.6   & 2.3   & 1.9   & 2.8   & 2.2   & 1.9   & 2.3   & 1.8   & 2.0   & 1.8   & 1.6   & 1.4   & 1.8 \\
\hline
\multicolumn{20}{l}{\textit{Additive term ($\widehat{t}^{add}/\widetilde{p}$, in \%)} }   \\
\hline
Mean  & 2.9   & 2.7   & 2.5   & 2.5   & 3.2   & 2.8   & 2.8   & 2.4   & 2.6   & 3.2   & 2.9   & 3.0   & 2.8   & 2.4   & 2.4   & 2.1   & 2.5   & 1.9   & 1.9   & 1.5 \\
Median & 2.0   & 1.8   & 1.6   & 1.3   & 2.0   & 2.0   & 2.2   & 1.6   & 2.0   & 2.5   & 1.9   & 2.2   & 1.9   & 1.8   & 2.1   & 1.7   & 1.9   & 1.6   & 1.6   & 0.8 \\
\hline
\# observations & 31865 & 32146 & 32344 & 33181 & 33986 & 34585 & 36090 & 36407 & 37255 & 37672 & 37757 & 41431 & 41763 & 39604 & 38950 & 37332 & 37748 & 38562 & 38387 & 38473 \\
\hline
\hline
\end{tabular}}

\end{center}
\label{tab_oa:result_vessel_ally3}%
\end{table}
\end{landscape}

\setcounter{table}{0}
\renewcommand{\thetable}{C.\arabic{table}}

\setcounter{figure}{0}
\renewcommand{\thefigure}{C.\arabic{figure}}

\subsection{Eliminating the composition effects: Primary vs. Manufacturing sector} \label{sec_oa:comp-effects}

In this section, we refine the characterization of the evolution of international transport costs by distinguishing primary goods trade flows and and manufactured goods trade flows. The evolution in transport costs over time, by transport mode (overall transport costs and composition effects excluded) are reported in Figure \ref{fig:totalTC_compeffects_excl_manuf} for the manufacturing sector, and in Figure \ref{fig:totalTC_compeffects_excl_primary} for the primary goods. For easing comparison, we also report the results obtained on the whole range of trade flows (i.e., Figure 2 of the paper), in Figure \ref{fig:totalTC_compeffects_excl}.

The classification retained to categorize trade flows follows the UNCTAD classification (on STIC Revision 3)\footnote{See "UNCTAD product groupings and composition (SITC Rev. 3)" in http://unctadstat.unctad.org/EN/Classifications.html, accessed Septembre 2018}.
Are considered as ``primary goods'' all flows recorded as ``Food and live animals'' (First digit ``0'' in the SITC Classification), ``Beverages and tobacco'' (First digit ``1''), ``Crude materials, inedible, except fuels'' (First digit ``2''), ``Mineral fuels, lubricants and related materials'' (First digit ``3''), ``Animal and vegetable oils, fats and waxes'' (First digit ``4''), ``Pearls, precious \& semi-precious stones'' (Classified ``667'' in the SITC Classification) and ``Non-ferrous metals'' (classified ``68'' in the SITC Classification).

\begin{figure}[htbp]
\caption{Transport costs (with and without composition effects), Manufacturing}
\label{fig:totalTC_compeffects_excl_manuf}
\begin{center}
\includegraphics[height=8cm] {graph_composition_manuf.pdf}
\end{center}
\end{figure}

\begin{figure}[htbp]
\caption{Transport costs (with and without composition effects), Primary goods}
\label{fig:totalTC_compeffects_excl_primary}
\begin{center}
\includegraphics[height=8cm]
{graph_composition_primary.pdf}
\end{center}
\end{figure}

\begin{figure}[htbp]
\caption{Transport costs (with and without composition effects)}
\label{fig:totalTC_compeffects_excl}
\begin{center}
\includegraphics[height=8cm]
{graph_composition_all.pdf}
\end{center}
\end{figure}

As reported in Figures \ref{fig:totalTC_compeffects_excl_manuf} and \ref{fig:totalTC_compeffects_excl_primary}, the ``raw'' transport costs and the unfitted transport costs, in the red dashed line plain have regularly declined over the period in both sectors, by roughly the same order of magnitude (50\% in air, 60\% in vessel). However the role of trade composition effects in accounting for this trend pattern differs depending on the sector.\\
Figure \ref{fig:totalTC_compeffects_excl_manuf} (for manufactures) reports a very similar time trend decomposition than what is obtained on the whole range of goods (Figure \ref{fig:totalTC_compeffects_excl}).  In air transport, most of the decrease can be imputed to the reduction of``ceteris paribus'' transport costs (the blue continuous line), trade composition effects playing virtually no role (Figure \ref{fig:totalTC_compeffects_excl_manuf}, left-hand panels (a), (b) and (c)). Trade composition effects matter more in vessel transport (Figure \ref{fig:totalTC_compeffects_excl_manuf}, right-hand panels (d), (e) and (f)), primarily in the ad-valorem component. As for the whole range of flows, the 60\% decrease in the unfitted transport costs in vessel can de decomposed in a 50\% decrease in the ``ceteris paribus'' transport costs (fitted), the 10\% remaining to trade composition effects.

The situation is is strikingly for primary goods only. In this case, it is in air transport that composition effects do matter (Figure \ref{fig:totalTC_compeffects_excl_primary}, left-hand panels (a), (b) and (c)), while we observe not much role for them in vessel transport (Figure \ref{fig:totalTC_compeffects_excl_primary}, left-hand panels (d), (e) and (f)). Furthermore, in air transport, composition effects matter by partially offsetting the decrease in the ``ceteris paribus'' transport costs (ie, implying a reduction in the ``raw'' transport costs over time much less pronounced than the fitted transport cost measures).

One explanation of these contrasted results lies in the share of primary goods in total flows. It is reported in Figure \ref{fig:Share_prim_goods}. In air transport, the share of primary goods in the total value of US imports is very small, around 10\% all over the period from 1974 to 2014. Primary goods make a higher proportion of trade flows in vessel transport, especially over 1974-1982 (between 40\% and 60\%). On the following sub-period though, their share has fallen to 20-30\%. Given the modest proportion of primary goods in total imports flows of the US economy compared to the manufactured sector, it is hence not surprising that the diagnosis made about the time trend of transport costs when all types of flows are considered is driven by the trend patterns that occur within the manufacturing sector. \\
Still, it is interesting to notice that composition effects are not dominated by changes in the primary goods share. If they were, we would not observe the same trends for trade as the whole and trade of manufactured goods.

\begin{figure}[htbp]
\caption{Share of primary goods in the value of total US imports}
\label{fig:Share_prim_goods}
\begin{center}
\includegraphics[height=6cm]
{Share_of_primary.pdf}
\end{center}
\end{figure}

\end{document}