### 2021-07-21 redaction/JEGeo/revision_JEGeo/referees/answers/general_answer_editor_revision.tex

\documentclass[12pt]{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%TCIDATA{OutputFilter=LATEX.DLL}
%TCIDATA{Version=5.50.0.2960}
%TCIDATA{<META NAME="SaveForMode" CONTENT="1">}
%TCIDATA{BibliographyScheme=Manual}
%TCIDATA{Created=Monday, February 14, 2011 11:40:51}
%TCIDATA{LastRevised=Thursday, February 17, 2011 16:51:46}
%TCIDATA{<META NAME="GraphicsSave" CONTENT="32">}
%TCIDATA{<META NAME="DocumentShell" CONTENT="Standard LaTeX\Blank - Standard LaTeX Article">}
%TCIDATA{CSTFile=40 LaTeX article.cst}

\usepackage{graphicx}
\usepackage{setspace}
%\usepackage{hyperref}
\usepackage{geometry}
\usepackage{blindtext}
\geometry{hmargin=3cm,vmargin=2.5cm}
\usepackage{lscape}
\newtheorem{theorem}{Theorem}
\newtheorem{acknowledgement}[theorem]{Acknowledgement}
\newtheorem{algorithm}[theorem]{Algorithm}
\newtheorem{axiom}[theorem]{Axiom}
\newtheorem{case}[theorem]{Case}
\newtheorem{claim}[theorem]{Claim}
\newtheorem{conclusion}[theorem]{Conclusion}
\newtheorem{condition}[theorem]{Condition}
\newtheorem{conjecture}[theorem]{Conjecture}
\newtheorem{corollary}[theorem]{Corollary}
\newtheorem{criterion}[theorem]{Criterion}
\newtheorem{definition}[theorem]{Definition}
\newtheorem{example}[theorem]{Example}
\newtheorem{exercise}[theorem]{Exercise}
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{notation}[theorem]{Notation}
\newtheorem{problem}[theorem]{Problem}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{remark}[theorem]{Remark}
\newtheorem{solution}[theorem]{Solution}
\newtheorem{summary}[theorem]{Summary}
\newenvironment{proof}[1][Proof]{\noindent\textbf{#1.} }{\ \rule{0.5em}{0.5em}}

\setlength{\parindent}{0pt}

\newcommand*\sepline{%
  \begin{center}
    \rule[1ex]{.5\textwidth}{.5pt}
  \end{center}}
\newcommand*\sepstars{%
  \begin{center}
    $\star\star\star$
  \end{center}}

\begin{document}

\noindent \textbf{Manuscript ID JOEG-2019-027, ``International Transport costs:\\New Findings from modeling additive costs'': revised revision}

\bigskip
\bigskip

\hfill Paris, \today

\bigskip
\noindent Dear Pr. Kerr, dear Bill, \\
\\
\noindent \noindent We would like to thank you for the opportunity to resubmit our paper, and for your patience in waiting the outcome. We are very grateful for the deadline extensions you kindly granted us. Obviously, the health crisis and the various consecutive lockdowns made the access to our offices difficult in several occurrences.

However, the major source of delays lies in our attempts to obtain additional data to accommodate a very important concern by Referee \#1 (see below). We ordered this new data from the Census Bureau at the end of December 2019. However, some of the years we requested (2000 and 2001, as well as years before 1998) are not on their servers: they are back at their office which they have not been able to access since last spring apparently, due to the current health situation. Since then, the situation has never evolved, and we are still currently waiting for these missing years. We can of course make available all our e-mail exchanges with Jeremy Sanchez, our contact at the Census Bureau. Until very recently, Mr. Sanchez never gave us any clear “yes or no” answer to our requests concerning a possible return in his DC office (closed because of the CoViD 19 restrictions), which would have allowed the delivery of the data. We can of course make available all our e-mail exchanges with Jeremy Sanchez. In his last e-mail, he confirmed that access to the additional data was impossible.

Considering the lack of credible perspective to access this additional data in a sufficiently close/foreseeable future, we completed the revision with the following data: a full dataset at the SITC5 level, completed with systematic robustness checks (see the list at the end of this e-mail) based on a HS10 dataset over the 2002-2019 period. These new estimates do not provide any substantial change to our benchmark estimates, and it seems reasonable to consider that additional checks including years from the 1980s and the 1990s would not dramatically change this outcome.

and to thank the referees for the comments which have helped us revise it. These suggestions have led to substantial improvements. We enclose detailed answers to the referees. Here we briefly list the major changes brought to the paper. We hope that you will be satisfied by the outcome.

\begin{enumerate}

\item Functional form/use of a non-linear estimator
\item Endogeneity issues/IV estimates
\item Aggregation issues
\item Calculation of unit prices
\item Big picture/usefulness of the exercise
\item XXX
\item XXX

\end{enumerate}

Altogether, we believe that the paper is much clearer and easier to read. The robustness exercises included in the main text focus on the elements that were the most problematic to the referees (XXX). We hope that you will be pleased by the outcome.

\bigskip

Yours sincerely,

\bigskip

\hfill Jérôme Héricourt

\hfill Professor of Economics

\hfill University of Lille (LEM-CNRS) and CEPII

\bigskip
\bigskip
\bigskip
\sepstars

\end{document}