### 2021-07-28 redaction/JEGeo/revision_JEGeo/revised_article/table_compWelfare2_air.tex
Continuer sur le modèle

\begin{tabular}{llllll}
& Mode (1 Air, 2 Ves) & Lower tau, same fx & Lower tau AND fx (export share = 35%) & Lower tau and t, same fx & Lower tau, t, AND fx (export share = 35%) \\ 
\hline 
TC(0) & 1 & 0.062 & 0.062 & 0.062 & 0.062 \\ 
TC(1) & 1 & 0.028 & 0.028 & 0.028 & 0.028 \\ 
Delta Total TC (pp) & 1 & -0.034 & -0.034 & -0.034 & -0.034 \\ 
Delta tau (pp) & 1 & -0.034 & -0.034 & -0.014 & -0.014 \\ 
Delta t/pfas (pp) & 1 & 0 & 0 & -0.020007 & -0.020007 \\ 
Delta Welfare (abs) & 1 & 0.020229 & 0.036246 & 0.029487 & 0.052655 \\ 
Rel. Delta welf (in %) & 1 & 1.2319 & 2.2073 & 1.8187 & 3.2477 \\ 
\hline 
\end{tabular}