### 2021-07-29 redaction/JEGeo/revision_JEGeo/revised_article/table_compWelfare2_vessel.tex

\begin{tabular}{llllll}
& Mode (1 Air, 2 Ves) & Lower tau, same fx & Lower tau AND fx (export share = 35%) & Lower tau and t, same fx & Lower tau, t, AND fx (export share = 35%) \\ 
\hline 
TC(0) & 2 & 0.105 & 0.105 & 0.105 & 0.105 \\ 
TC(1) & 2 & 0.041 & 0.041 & 0.041 & 0.041 \\ 
Delta Total TC (pp) & 2 & -0.064 & -0.064 & -0.064 & -0.064 \\ 
Delta tau (pp) & 2 & -0.064 & -0.064 & -0.031 & -0.031 \\ 
Delta t/pfas (pp) & 2 & 0 & 0 & -0.032947 & -0.032947 \\ 
Delta Welfare (abs) & 2 & 0.035413 & 0.046674 & 0.052202 & 0.064189 \\ 
Rel. Delta welf (in %) & 2 & 2.1791 & 2.8719 & 3.2951 & 4.0517 \\ 
\hline 
\end{tabular}