### 2021-09-09 redaction/JEGeo/revision_JEGeo/revised_article/donnees_calibration.tex
changement .tex sur le modele

\begin{tabular}{lll}
& Value & Reference\\
\hline
$\sigma$ & 4 & \cite{Irrazabal_2015}\\
$f_e$ & 1 & \cite{ghironi}\\
$f$ & 0.033 & \cite{ghironi} \\
$L$ & 1 & Arbitrary\\
$\delta$ & 0.1 & \cite{ghironi}, \cite{Irrazabal_2015}\\
$k$ & 4 \\
$\varphi_{\text{min}}$ & 1 & \cite{ghironi}, \cite{Irrazabal_2015}\\ \hline
\multicolumn{3}{l}{Fixed export cost } \\
Share of exporting firms & 21\% &  \cite{ghironi}, based on BEJK, 2003  \\
Share of exporting firms & 39\% &  \cite{Lincoln_McCallum2018} \\
\hline
\end{tabular}