### 2021-09-21 redaction/JEGeo/revision_JEGeo/revised_article/donnees_calibration.tex

\begin{tabular}{lll}
& Value & Reference\\
\hline
$\sigma$ & 4 & \cite{Irrazabal_2015}\\
$f_e$ & 1 & \cite{ghironi}\\
$f$ & 0.033 & \cite{ghironi} \\
$L$ & 1 & Arbitrary\\
$\delta$ & 0.1 & \cite{ghironi}, \cite{Irrazabal_2015}\\
$k$ & 4 \\
$\varphi_{\text{min}}$ & 1 & \cite{ghironi}, \cite{Irrazabal_2015}\\ \hline
\multicolumn{3}{l}{Fixed export cost } \\
$f_x $ & 0.08 & Share of exporting plants = 21\% (\cite{ghironi}, based on BEJK, 2003) \\
& 0.06& Share of exporting plants = 39\% (\cite{Lincoln_McCallum2018}) \\
\hline
\end{tabular}